---
title: "BNB-Price-Prediction-Project"
author: "Shrey Jaradi"
date: "2022-11-23"
output: html_document
---

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(psych)
library(ggmap)
library(GGally)
library(caret)
library(glmnet)
library(randomForest)
library(e1071)
library(lmvar)
library(maps)
library(rpart)
library(rpart.plot)
cat("Library loaded Successfully")
```

```{r}
bnb_df = read.csv(file='../data/AB_NYC_2019.csv') 
head(bnb_df)
```
```{r}
dim(bnb_df)
```
```{r}
summary.default(bnb_df)
```

**name :** Name of the Place 
**host_name :** It's a host name - name of owner  - not much important          
**neighbourhood_group :** This also we can convert to dummy variable -as this is also important- there are certain negihbourhood where price is varies  
**neighbourhood :** There are so many values if we create dummy variable for this, it's gonna create so much feature better to ignore this column            
**Room Type :**  we can convert to Dummy variable as there are three Room Type and it will help us as this is much relevant column 
**last_review  :** This is date column we can convert it to datetime column. 

```{r}
cat("Room Type : ", length(unique(bnb_df$room_type)))
cat("\nNeighbourhood group : ",length(unique(bnb_df$neighbourhood_group)))
cat("\nNeighbourhood : ",length(unique(bnb_df$neighbourhood)))
```

There are 1052 missing values in "reviews_per_month" and "last_review" column,
as this are the column for datetime and number of reviews  last_review, 
we can replace it with NA for last_review and 0 for reviews_per_month.

```{r}
naniar::gg_miss_var(bnb_df) +
  theme_minimal()+
  labs(y = "Look at all the Missing Values") 
```

```{r}
bnb_df['last_review'][is.na(bnb_df['last_review'])] = 'NA'
bnb_df['reviews_per_month'][is.na(bnb_df['reviews_per_month'])] = 0
bnb_df = bnb_df[(bnb_df$price!=0),]
```

```{r}
naniar::gg_miss_var(bnb_df) +
  theme_minimal()+
  labs(y = "Look at all the Missing Values") 
```

```{r}
skewness(bnb_df$price) # > 1 = highly skewed towards the right
less_price_bnb = bnb_df %>% filter(price < 1000)
skewness(less_price_bnb$price) # > 1 but much less skewed than before
```
 
```{r}
par(mfrow=c(1,2))
qqnorm(less_price_bnb$price)
qqline(less_price_bnb$price)
qqnorm(bnb_df$price)
qqline(bnb_df$price)
```

Outlier Identification 
```{r}
Quant = quantile(bnb_df$price, probs=c(.25, .75), na.rm = T)
iqr_val = IQR(bnb_df$price, na.rm = T)
bnb_df_new = bnb_df %>% filter(price > (Quant[1] - 1.5*iqr_val) & 
                       price < (Quant[2] + 1.5*iqr_val))

```
 
Outlier Visualization 
```{r}
par(mfrow=c(2,1))
options(repr.plot.width=12, repr.plot.height=6)
boxplot(bnb_df_new$price, col = "yellow", horizontal = T, main = "After Removing Outliers - Price")
boxplot(bnb_df$price, col = "green", horizontal = T, main = "Before Removing Outliers Price")
```
 
Let's try to understand the statistics by Room Type , As we are trying to predict price , Let's check what are the minimum, maximum, average price according to room type and neighbourhood group

Manhattan is costliest area, their average price is almoot $250.
```{r}
options(dplyr.summarise.inform = FALSE)
bnb_df_new %>% group_by(room_type, neighbourhood_group) %>% dplyr::summarise(min_price = min(price), avg_price = mean(price),max_price = max(price))
```
Price in Neighbourhood_group by room Type 

```{r}
ggplot(bnb_df_new, aes(x = neighbourhood_group, y = price, fill = room_type)) +
  geom_bar(position = "dodge", stat = "identity") +
  theme_classic() +
  labs(title = "NYC Airbnb Neighbourhood vs Price", x = "Neighbourhood", y = "Price") +
  theme(plot.title = element_text(face = "bold"))
```
Rental Price Distribution
```{r}
pal = c("olivedrab3", 'yellow', 'orange', 'grey50')
options(repr.plot.width=14, repr.plot.height=7)
ggplot(bnb_df_new, aes(price)) + 
  geom_density(show.legend = F, lwd = 1.5, fill = "yellow") + 
  scale_fill_manual(values=pal) + theme_bw(base_size = 10)  +
  labs(title = "Price Disribution", x = "Price")
```

Costliest Neighbourhood 
```{r}
grp_neighbourhood_df = bnb_df_new %>% 
  group_by(neighbourhood_group, price) %>% 
  tally(sort = TRUE) %>% 
  filter(row_number() <= 5)
options(repr.plot.width=14, repr.plot.height=7)
ggplot(data = grp_neighbourhood_df, aes(x = neighbourhood_group, y = price)) +
  geom_col(fill="lightblue", lwd=1.5) + 
scale_fill_manual(values=pal) + theme_bw(base_size = 10)
```
Top 10 Neighbourhood 
```{r}
bnb_df_new %>% group_by(neighbourhood) %>%
  dplyr::summarize(num_listings = n(), borough = unique(neighbourhood_group)) %>%
  top_n(n = 10, wt = num_listings) %>% ggplot(aes(x = fct_reorder(neighbourhood, num_listings),y = num_listings, fill = borough)) +geom_col() + coord_flip() +theme(legend.position = "bottom") + labs(title = "Top 10 neighborhoods", x = "Neighborhood", y = "Number of listings")
```



Import NYC Map and show datapoint , according to price, which are costly neighboorhood and preferrred one
```{r}
bnb_df %>% group_by(neighbourhood_group) %>% dplyr::summarise(avg_price = mean(price))

bnb_df = bnb_df %>%
  mutate(avg_price = case_when(
    endsWith(neighbourhood_group, "Bronx") ~ 87.57706,
    endsWith(neighbourhood_group, "Brooklyn") ~ 124.43892,
    endsWith(neighbourhood_group, "Manhattan") ~ 196.88490,
    endsWith(neighbourhood_group, "Queens") ~ 99.51765,
    endsWith(neighbourhood_group, "Staten Island") ~ 114.81233
    ))
```


Price range within Neighbourhood
```{r}
height = max(bnb_df$latitude) - min(bnb_df$latitude)
width = max(bnb_df$longitude) - min(bnb_df$longitude)
LA_borders = c(bottom  = min(bnb_df$latitude)  - 0.1 * height, 
                top     = max(bnb_df$latitude)  + 0.1 * height,
                left    = min(bnb_df$longitude) - 0.1 * width,
                right   = max(bnb_df$longitude) + 0.1 * width)

map = get_stamenmap(LA_borders, zoom = 10, maptype = "toner-lite")
ggmap(map) + geom_point(data = bnb_df, mapping = aes(x = longitude, y = latitude, 
                                               col = log(price))) + scale_color_distiller(palette = "RdYlGn", direction = 1)
```

```{r}
states = map_data("state")
new_york = subset(states, region %in% c("new york"))
ggplot(data = new_york) +
    geom_polygon(aes(x = long, y = lat, group = group), fill = "gray", color = "white") + 
    geom_point(data=bnb_df_new, aes(x=longitude, y=latitude, color = neighbourhood_group), alpha=0.3) +
    coord_fixed(xlim = c(-74.4, -73.5),  ylim = c(40.3, 41), ratio = 1.2) + 
    scale_fill_gradient(low = "blue", high = "red")
```

Relationship between price and num of reviews 
```{r}
ggplot(bnb_df_new, aes(number_of_reviews, price)) + theme(axis.title = element_text(), axis.title.x = element_text()) +
geom_point(aes(size = price), alpha = 0.05, color = "blue") + xlab("No. of reviews") + ylab("Price") + ggtitle("Relationship between number of reviews", subtitle = "Expensive objects have small no. of reviews")
```


Differences in Means that we need to account for an interaction term o in our Regression model
```{r}
par(mfcol=c(1,2))
boxplot(price ~ neighbourhood_group, data = bnb_df_new, frame = F, col = "lightgreen")
boxplot(price ~ room_type, data = bnb_df_new, frame = F, col = "lightblue")
```


Need to convert the categorical data to level 
```{r}
bnb_df_new$neighbourhood = as.factor(bnb_df_new$neighbourhood)
bnb_df_new$neighbourhood_group = as.factor(bnb_df_new$neighbourhood_group)
bnb_df_new$room_type = as.factor(bnb_df_new$room_type)
bnb_df_new$last_review = as.POSIXct(bnb_df_new$last_review, format="%Y-%m-%d", tz="UTC")

```

```{r}
summary.default(bnb_df_new)
```

```{r}
head(bnb_df_new)
```

Filter the Categorical Columns, to get the correlation of matrix
```{r}
filter_cat_bnb_df = bnb_df_new [,c("id","host_id","neighbourhood_group","neighbourhood", "latitude", "longitude", "room_type", "price", "minimum_nights", "number_of_reviews" ,"last_review", "reviews_per_month", "calculated_host_listings_count","availability_365")] 
```


```{r}
corr_bnb_df =  bnb_df_new [,c("neighbourhood_group", "room_type", "price", "minimum_nights", "number_of_reviews" ,"reviews_per_month", "calculated_host_listings_count","availability_365")] 
corr_bnb_df = corr_bnb_df[, sapply(corr_bnb_df, is.numeric)]
corr_bnb_df = corr_bnb_df[complete.cases(corr_bnb_df),]
```

```{r}
pairs.panels(corr_bnb_df, method = "pearson", hist.col = "#00AFBB",density = TRUE,ellipses = TRUE)
```
####Machine Learning - Linear Regression

Data Split into train and test !!
```{r}
bnb_df_index = createDataPartition(filter_cat_bnb_df$price, p =.70,list=FALSE)
bnb_df_train =  filter_cat_bnb_df[bnb_df_index,] 
bnb_df_test =   filter_cat_bnb_df[-bnb_df_index,] 
```

Excluding Features Reasons : 

- id: not relevant because it's unique number
- name: not relevant because it's  Identifier
- host_id: not relevant because it's unique number 
- host_name: not relevant because it's  Identifier
- neighbourhood: as we already take neighbourhood_group in our study, it's a redudant feature
- last_review: datetime variable, unnecessarily will complicate our model

```{r}
bnb_model = lm(price ~ neighbourhood_group + scale(latitude) +  scale(longitude) + room_type+ minimum_nights + number_of_reviews + reviews_per_month + calculated_host_listings_count + availability_365 , data=bnb_df_train)
summary(bnb_model)

```

The $R^2$ values is two low for this model, we can transform our Price feature and 
```{r}
par(mfrow=c(2,2))
plot(bnb_model)
```
Doing the Logarithmic Transformation on Y(Price) Predicted feature and 
Scaling the Latitude and Longitude Variable, 
```{r}
bnb_model_two = lm(log(price) ~   neighbourhood_group + scale(latitude)+  scale(longitude) + room_type+ minimum_nights + number_of_reviews + reviews_per_month + calculated_host_listings_count + availability_365 , data=bnb_df_train)
summary(bnb_model_two)
```

The 2nd model Q-Q plot, is much better than the previous model. We can clearly see that, and if you see the $R^2$ values also seems to be better than our previous model.

```{r}
par(mfrow=c(2,2))
plot(bnb_model_two)
```

```{r}
y_bnb_df_test = bnb_df_test[,c("neighbourhood_group", "latitude", "longitude", "room_type", "minimum_nights", "number_of_reviews" , "reviews_per_month", "calculated_host_listings_count","availability_365")] 
predicted_val = predict(bnb_model_two, y_bnb_df_test)
predicted_val = exp(predicted_val)
```

```{r}
RMSE_regression = sqrt(mean( (bnb_df_test$price - predicted_val)**2 ))
SSE = sum((bnb_df_test$price - predicted_val)**2)
SSR = sum((predicted_val - mean(bnb_df_test$price)) ** 2)
R2 = 1 - SSE/(SSE + SSR)
cat("Linear Regression RMSE on Test Dataset : ",RMSE_regression)
cat("\nLinear Regression  SSE on Test Dataset  : ",SSE)
cat("\nLinear Regression  SSR on Test Dataset  : ",SSR)
cat("\nLinear Regression  R-Square  on Test Dataset  : ",R2)
```

A Simple Linear Regression technique using only neighbourhood as predictor variable
```{r}
bnb_model_three = lm(log(price) ~   neighbourhood, data=bnb_df_train)
summary(bnb_model_three)
```


```{r}
bnb_model_four = lm(log(price) ~  room_type, data=bnb_df_train)
summary(bnb_model_four)
```

```{r}
bnb_model_five = lm(log(price) ~  neighbourhood_group, data=bnb_df_train)
summary(bnb_model_five)
```

Predict price by different neighbourhood, room type, availability_365, minimum_nights + number of reviews 
```{r}
bnb_model_six = lm(log(price) ~  neighbourhood + room_type + availability_365 + minimum_nights + number_of_reviews , data=bnb_df_train)
summary(bnb_model_six)
```

#### Decision Tree Regressor

```{r}
dt_bnb_model = rpart(formula = price ~ neighbourhood_group + latitude+  longitude + room_type+ minimum_nights + number_of_reviews + reviews_per_month + calculated_host_listings_count + availability_365, data = bnb_df_train,method = "anova")
printcp(dt_bnb_model)
```

```{r}
prp(dt_bnb_model, main = "Decision Tree Model")
```

```{r}
predict_dt_val = predict(dt_bnb_model, bnb_df_test)
```



```{r}
MSE_dt = sum((predict_dt_val - bnb_df_test$price)^2)/nrow(bnb_df_test)
vary_dt = sum((bnb_df_test$price - mean(bnb_df_test$price))^2)/(nrow(bnb_df_test)-1)
Rsqr_dt = 1 - (MSE_dt/vary_dt)
MSE_dt
vary_dt
Rsqr_dt
```

### Random Forest
```{r}
set.seed(4543)
rf_model = randomForest(price ~  neighbourhood_group + latitude+  longitude + room_type+ minimum_nights + number_of_reviews + reviews_per_month + calculated_host_listings_count + availability_365, data=bnb_df_train, na.action = na.omit)
rf_model
```

```{r}
plot(rf_model)
```

```{r}
predict_rf_val = predict(rf_model, bnb_df_test)
```

```{r}
MSE_rf = sum((predict_rf_val - bnb_df_test$price)^2)/nrow(bnb_df_test)
vary_rf = sum((bnb_df_test$price - mean(bnb_df_test$price))^2)/(nrow(bnb_df_test)-1)
Rsqr_rf = 1 - (MSE_rf/vary_rf)

cat("Random Forest MSE on Test Dataset : ",MSE_rf)
cat("\nRandom Forest variance on Test Dataset  : ",vary_rf)
cat("\nRandom Forest R-Square  on Test Dataset  : ",Rsqr_rf)
```


```{r}
varImpPlot(rf_model) 
```


